Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-09-23 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

\subsection{Influence of a moving beam front on seed crystals}
\textcolor{red}{Need to write up these results. To summarise the crystal 
	dissolved around the laser focus instead of growing as the seed crystals 
	are in the solution and not close to the air-solution interface}

>>>>

\subsection{Influence of a moving beam front on seed crystals}
One potential application of this phenomena would be in using a 
moving the moving beam front as a method for shaping the final
crystal morphology. Glycine seed crystals were grown via 
evaporative crystallisation over night. The resultant crystals
were as wide as $0.5 mm$ in some cases. Individual crystals were 
collected and suspended in a water + glycine solution ($S = 1.001$)
and the laser was switched on.

### 2024-09-23 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

\subsection{Dimer}
The dimer is defined by two spheres with refractive indices
of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
requires a target origin for setting the scattering expansion, as
with our simulations we set it as the dimer's centre of diffusion.
The near field scattering from the dimer is shown in 
fig.~\ref{fig:nf_scattering}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{nf_scattering_dimer.png}
	\caption{Re ${\bf E_z}$ distribution from a symmetric dimer when 
	irradiated by the probe beam (viewed in the x-y plane). Axis are 
	given in dimensionless units of $k\hat{x}$ and $k\hat{y}$ respectively.}
	\label{fig:nf_scattering}
\end{figure}

\subsection{Detectors and Pixels}
Each detector is placed some distance $\bf r$ from the origin, 
detector positions are defined by angles $\theta$ \& $\phi$. With 
$\theta=\phi=0$ being directly in front of the probe beam. The 
position of each detector is given by:
\begin{align}
	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
\end{align}
Where r is the distance between the detector and the target origin. 
In an ideal situation the surface of the detector is oriented towards 
the target origin perfectly. Pixels can be thought of as points lying 
on the surface of the detector, each point has its own position in real
space. Using \textit{mstm} we can find the components of the scattering 
matrix at each point and compute the intensity of the electric field via:

\begin{align}
	\begin{pmatrix}
		I_s \\ Q_s \\ U_s \\ V_s
	\end{pmatrix} = 
	\begin{pmatrix}
		S_{11} & S_{12} & S_{13} & S_{14} \\
		S_{21} & S_{22} & S_{23} & S_{24} \\
		S_{31} & S_{32} & S_{33} & S_{34} \\
		S_{41} & S_{42} & S_{43} & S_{44} \\
	\end{pmatrix}
	\begin{pmatrix}
		I_i \\ Q_i \\ U_i \\ V_i
	\end{pmatrix}
\end{align}
Where:
\begin{align}
	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2 \\
	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
\end{align}

By computing the scattering at each point on the detectors surface 
we can compute the average scattering over the detector. As the 
detectors are moved further away from the target origin the intensity 
distribution across the detector surface falls off until near uniform.

By moving the detectors further back we can treat each detector as single
point and compute the intensity at the centre of the detector rather than
consider the full surface.

\section{Interpretation of scattering data into orientation estimates}
Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any 
point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere 
to the centre of the smaller sphere. A plane wave probe beam is incident on 
the trapping laser, is incident on the dimer, generating a scattering pattern
dependent on the dimer's orientation $I(\hat{\bf s}, \theta, \phi)$ which can be 
computed using \textit{mstm} \cite{I.Mishchenko1996}. To represent the experimental 
set up consisting of a set of optical fibres recording scattered light, we choose 
four sets of spherical angles [$(\theta_1,\ \phi_1), \ (\theta_2,\ \phi_2), \ 
(\theta_3, \ \phi_3), \ (\theta_4,\ \phi_4)$] and record the calculated intensity 
at each angle $I(\hat{s},\ \theta_k,\ \phi_k)$.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

Our goal is to determine the orientation of the trapped dimer based on the measured 
intensity $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate 
of the dimer's orientation, for the purposes of interpretation of the scattering and 
optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then 
use as 'classification categories' in a neural network methodology to map scattering 
data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} 
\ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a 
unit sphere \cite{Reyuthor2006} (Figure~\ref{fig:dimer}b) leading to a maximum 
nearest-neighbour spacing between two neighbouring reference orientations of 0.895 
radians. Using MSTM we compute the raw intensities at each of the measurement angles 
that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary 
there are several constraining factors that limit our ability to infer useful information 
from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown 
of our choice of detection angles. The raw intensities are normalized according to:
\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_\alpha)
	= 
	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}

where the denominator is simply the standard deviation across the set of values
$I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled 
signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to 
their associated reference orientations: as is well known from such examples of 
the inverse scattering problem. While it is trivial to compute the light scattering 
pattern for any given particle with any particular characteristic (i.e. size, 
shape, or orientation), inferring the light scattering from a unknown particle 
to determine said characteristic is incredibly difficult due to complex mapping 
between scattering and said characteristic. Even if the orientation space is divided 
evenly between reference orientation the subsequent signal space ends up being 
appearing mixed making simple comparisons of signals useless for inferring information 
on the particle. Shown below is two clusters of orientation vectors and there 
respective measured scattering signals - the points have been coloured based on 
their proximity to the centre of their respective cluster. While the orientation 
space appears tightly packed and ordered the signal space quickly spreads out in 
an asymmetric fashion. Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
mapping can intersect itself which only further increases the complexity. While 
in some instances the mapping between one reference orientation and another is 
discrete, in other instances the mapping becomes far more complex to discern.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

Nevertheless, at least where the uncertainty in signal measurements is low (see 
below), we can predict the orientation from the scattering by utilising 
computational techniques such as neural networks. We thus utilised the Python 
machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a 
neural network for identifying the dimer's orientation from its light scattering 
signal. The network was trained by generating a database of random orientation 
vectors, calculating the corresponding light scattering signals, and then using 
the network to estimate the probability of a given signal coming from a dimer in 
a given reference orientation. The network's loss function was evaluated and used 
to improve the estimation, the network being trained until the improvement in the 
loss function was less than 0.0001.

@@

where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of 
the distributions of particle orientations and instantaneous signals, respectively.
\textit{Without} any prior evidence we must assume that the orientation prior of 
the dimer $p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
dimer's possible current orientation from knowledge of previous measurements can 
be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$. The latter prior 
$p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$). This is 
given by taking the discrete integral over the collection of reference orientations:
\begin{align}
	p(y_1, y_2, y_3, y_4)
	=
	\sum_{\alpha=1}^{n_{\rm ref}}
	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
	p(\hat{\bf n}_\alpha)
\end{align}

@@

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$. We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}. We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.

@@

One reason we observe such large jumps in orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space: even a large change in orientation can involve a small change in scattering. Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step. To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior, accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness. Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step as a Boltzmann distribution of the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are reweighing our estimation based on the size of rotation required, with smaller movements being favoured over large movements:

\begin{align}
	p(\hat{\bf n}_\alpha)
	&=\frac{e^{\beta (\hat{\bf n}_\alpha 
			\cdot \hat{\bf n}_{est}(t-\Delta t))}}
	{\sum_{\alpha=1}^{n_{\rm ref}}
		e^{\beta (\hat{\bf n}_\alpha 
			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
	\label{eq:boltz}
\end{align}
Here $\beta$ is a weighting factor describing the dimer's freedom of motion within the trap. As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.

>>>>

\subsection{Dimer}
The dimer is defined by two spheres with refractive indices
of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
requires a target origin for setting the scattering expansion, as
with our simulations we set it as the dimer's centre of diffusion.

\subsection{Scattering and Detection}
Each detector is placed some distance $\bf r$ from the origin, 
detector positions are defined by angles $\theta$ \& $\phi$. With 
$\theta=\phi=0$ being directly in front of the probe beam. The 
position of each detector is given by:
\begin{align}
	\begin{pmatrix}
		x_{fiber} \\ y_{fiber} \\ z_{fiber}
	\end{pmatrix} = 
	\begin{pmatrix}
		rcos(\phi)sin(\theta) \\ rsin(\phi)sin(\theta) \\ rcos(\theta)]
	\end{pmatrix}
\end{align}
Where r is the distance between the detector and the target origin. 
In an ideal situation the surface of the detector is oriented towards 
the target origin perfectly. Pixels can be thought of as points lying 
on the surface of the detector, each point has its own position in real
space. Using \textit{mstm} we can find the components of the scattering 
matrix at each point and compute the intensity of the electric field via:

\begin{align}
	\begin{pmatrix}
		I_s \\ Q_s \\ U_s \\ V_s
	\end{pmatrix} = 
	\begin{pmatrix}
		S_{11} & S_{12} & S_{13} & S_{14} \\
		S_{21} & S_{22} & S_{23} & S_{24} \\
		S_{31} & S_{32} & S_{33} & S_{34} \\
		S_{41} & S_{42} & S_{43} & S_{44} \\
	\end{pmatrix}
	\begin{pmatrix}
		I_i \\ Q_i \\ U_i \\ V_i
	\end{pmatrix}
\end{align}
Where:
\begin{align}
	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2  \label{eq:intensity} 
	\\
	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
\end{align}

By computing the scattering at each point on the detectors surface 
we can compute the average scattering over the detector. From a
experimental perspective the distance from the particle to the 
detectors must be considered. The near field scattering can be 
computed with relative ease using \textit{mstm} as shown in 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{near_field_intensity_side_profile.png}
	\caption{Intensity distribution (as defined by \eqref{eq:intensity}) 
		from a symmetric dimer when irradiated by the probe beam 
		propagating in the $k\hat{z}$ direction (viewed in the 
		z-y plane). Axis are given in dimensionless units of 
		$k\hat{z}$ and $k\hat{y}$.}
	\label{fig:nf_scattering}
\end{figure}

While \textit{mstm} does provide the amplitude matrix it is not
scaled to account for distance from the target to the detector.
We amend this by simply scaling according to the inverse square 
law. As the detector is moved further from the target the 
intensity distribution across the entire detector will decrease
substantially. Because of the fall off it is more computationally 7
efficient to instead consider only the central pixel of the 
detector rather than across the entire surface. If we consider 
the difference between the average intensity and the intensity 
of the central point we can see that after not even $20 \mu$ 
away from the target particle the average intensity across the 
surface is within $0.1\%$ of the central pixel. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
	\caption{Normalised difference between the central pixel intensity
	$I(\theta, \phi)$ and the average pixel intensity over the surface
	of the optical fibre $\langle I \rangle$ as a function of distance.
	4 different optical detectors were tested (spherical coordinates 
	are given in the legend). While the choice of coordinates is 
	arbitrary they all demonstrate the same exponential decay.}
\end{figure}

By moving the detectors further back we can treat each detector 
as single point and compute the intensity at the centre of the 
detector rather than consider the full surface. For computing 
the instantaneous orientation we only consider the intensity 
at the centre of the detector.

\section{Interpretation of scattering data into orientation estimates}
We now discuss the methodology for computing the instantaneous 
orientation of an optically bound particle based on scattering
detected by a finite grouping of optical fibres.

Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), 
we can define at any point in time a unit vector $\hat{s}$ 
pointing from the centre of the larger sphere to the centre 
of the smaller sphere. A plane wave probe beam incident, 
is incident on the dimer, generating a scattering pattern
dependent on the dimer's orientation $I(\hat{\bf s}, 
\theta, \phi)$ which is computed using \textit{mstm} 
\cite{I.Mishchenko1996}. To represent the experimental set 
up consisting of a set of optical fibres recording scattered 
light, we choose four sets of spherical angles [$(\theta_1,\ 
\phi_1), \ (\theta_2,\ \phi_2), \ (\theta_3, \ \phi_3), \ 
(\theta_4,\ \phi_4)$] and record the calculated intensity at 
each angle $I(\hat{s},\ \theta_k,\ \phi_k)$. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

Our goal is to determine the orientation of the trapped dimer 
based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
Rather than aim an exact estimate of the dimer's orientation, 
for the purposes of interpretation of the scattering and 
optimisation of the measurement setup it is more convenient 
to discretize the possible orientation space into a number 
of possible reference orientations. We can then use each as 
'classification categories' in a neural network methodology 
to map scattering data to orientation (see below for further 
discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
reference orientations $\hat{\bf n}_{\alpha}$  evenly 
distributed on a unit sphere \cite{Reyuthor2006} 
(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
neighbour spacing between two neighbouring reference 
orientations of 0.895 radians. Using \textit{mstm} we 
compute the raw intensities at each of the measurement angles 
that would be generated by a dimer in each reference 
orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the 
number and position of detection fibres is technically arbitrary 
there are several constraining factors that limit our ability 
to infer useful information from the trapped object, see Section~\ref{sec:brownian} for a detailed breakdown of our 
choice of detection angles. In order for a neural network
to properly weight each signal the raw intensities are 
normalized according to:
\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_\alpha)
	= 
	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}

where the denominator is simply the standard deviation across 
the set of values $I(\hat{\bf n}, \theta_k)$. Note that the 
collected scattering signals are not necessarily simply 
related to their associated reference orientations: as is well 
known from such examples of the inverse scattering problem. 
While it is trivial to compute the light scattering pattern 
for any given particle with any particular characteristic (i.e. 
size, shape, or orientation), inferring the light scattering 
from a unknown particle to determine said characteristic is 
incredibly difficult due to complex mapping between scattering 
and said characteristic. Even if the orientation space is 
divided evenly between reference orientations the subsequent 
signal space ends up being appearing mixed making simple 
comparisons of signals useless for inferring information on the 
particle.

Fig~\ref{fig:mixing} shows two clusters of orientation 
vectors and their respective signals collected from 3 detectors
- the points have been coloured based on their proximity 
to the centre of their respective cluster. While the 
orientation space appears tightly packed and ordered the 
signal space quickly spreads out in an asymmetric fashion. 
Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
mapping can intersect itself which only further increases 
the complexity. While in some instances the mapping between 
one reference orientation and another is discrete, in other 
instances the mapping becomes far more complex to discern.
\newpage 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.8\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.8\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}
\newpage
Nevertheless, at least where the uncertainty in signal measurements 
is low, we can predict the orientation from the scattering scattering 
by utilising computational techniques such as neural networks. We 
thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
dimer's orientation from its light scattering signal. The network 
was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then 
using the network to estimate the probability of a given signal coming 
from a dimer in a given reference orientation. The network's loss 
function was evaluated and used to improve the estimation, the 
network being trained until the improvement in the loss function 
was less than 0.0001.

@@

where $p(\hat{\bf n}_\alpha)$ and $p(y_k(\hat{\bf s}))$ are the prior 
estimates of the distributions of particle orientations and 
instantaneous signals, respectively. \textit{Without} any prior 
evidence we must assume that the orientation prior of the dimer 
$p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
dimer's possible current orientation from knowledge of previous 
measurements can be used to inform our estimate of $p(\hat{\bf 
n}_{\alpha})$. The latter prior $p(y_k(\hat{s}))$ is the probability of 
measuring a signal ($y_1$, $y_2$, $y_3$). This is given by taking 
the discrete integral over the collection of reference orientations:
\begin{align}
	p(y_k(\hat{\bf s}))
	=
	\sum_{\alpha=1}^{n_{\rm ref}}
	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
	p(\hat{\bf n}_\alpha)
\end{align}

@@

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mackowski1996} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$. We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}. We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.

@@

One reason we observe such large jumps in the estimated 
orientation is simply because the neural network is assuming
that the signals are unrelated from one another. As demonstrated
by fig~\ref{fig:mixing} the orientation and signal spaces are 
uncorrelated with one another which results in the neural network
to assign orientations incorrectly. Combining this fact with 
use of a uniform prior ($p(\hat{\bf n}_\alpha)$), there 
is no constraint on how much estimated orientation can change 
from time-step to time-step.

To improve the estimation we can therefore use knowledge of the 
physical limitations of the object in the trap and its dynamics. 
Imposing a more physically grounded prior, accounting in this case 
for the fact that the rotational motion of the dimer is limited by 
the rotational trap stiffness ($\kappa_r$). Here the prior of the 
reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at 
each time step as a Boltzmann distribution of the physical distance 
between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and 
each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are 
reweighing our estimation based on the size of rotation, with smaller movements being favoured over large movements:
\begin{align}
	p(\hat{\bf n}_\alpha)
	&=\frac{e^{\hat{\bf n}_\alpha 
			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
	{\sum_{\alpha=1}^{n_{\rm ref}}
		e^{\hat{\bf n}_\alpha 
			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
	\label{eq:boltz}
\end{align}

As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.
